---
title: "HW 6: Linear Models"
author: "Emma Gorin"
date: "December 2020"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```

## Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide_df =
  read_csv("data/homicide-data.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1,
    )
  ) %>% 
  filter(city_state != "Tulsa, AL",
         victim_race %in% c("White", "Black")) %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

glm(resolution ~ victim_age + victim_race + victim_sex,
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr:: kable(digits = 3)
```


Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing Black victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

In most cities, odds of homicide being resolved by arrest are lower in male vs female victims.

## Problem 2

Class notes:
build a model of our choice, compare to two given models using cross validation.
If you give R an interaction term only, it will look at the main effects as well automatically.
Approach: Maybe start with given models and compare using cross validation, then look at new model and add it to the other results. If my model doesn't work as well as the first one, maybe reconsider.
Example of looking at residuals:
```{r}
#First make model called model_fit
baby_df %>% 
  modelr::add_residuals(model_fit) %>% 
  ggplot(aes(x = gaweeks, y = resid)) +
  geom_point()
```


In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
birthweight_df =
  read_csv("./data/birthweight.csv") %>% 
  mutate(
    baby_sex = recode(babysex, `1` = "male", `2` = "female")
  ) %>% 
  rename(
    ga_weeks = gaweeks,
    baby_length_cm = blength,
    baby_head_cm = bhead,
    baby_weight_g = bwt,
    mom_age = momage,
    mom_weight_lb = delwt,
    smoke_n = smoken
  ) %>% 
  select(baby_weight_g, baby_length_cm, ga_weeks, baby_head_cm, baby_sex, mom_age, mom_weight_lb, smoke_n) # to add: anything I use for a model
```


Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

**I was curious about the potential association between maternal smoking and birthweight. In addition to smoking, I decided to include maternal age and maternal weight in the model, which I think could be potential confounders. I also included gestational age and length at birth, since I was curious if maternal smoking would be associated with low birth weight even after adjusting for these factors.**

```{r}
my_mod = lm(baby_weight_g ~ mom_age + mom_weight_lb + ga_weeks + baby_length_cm + smoke_n, data = birthweight_df)
birthweight_df %>% 
  modelr::add_residuals(my_mod) %>% 
  modelr::add_predictions(my_mod) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```

**As you can see in the plot above, there is one baby with a predicted birthweight of < 0. I investigate this further below.**

```{r}
birthweight_df %>% 
  modelr::add_residuals(my_mod) %>% 
  modelr::add_predictions(my_mod) %>% 
  filter(pred < 1000) %>% 
  head()
```

**It turns out this particular baby's length in cm is listed as 20 - which seems implausible based on its gestational age. I decided to remove this observation from the dataset for the rest of the problem. **

```{r}
birthweight_df = 
  birthweight_df %>% 
  filter(baby_length_cm > 20)
```


Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
cv_df = 
  crossv_mc(birthweight_df, 100) %>%
  mutate(
    length_ga_mod = map(train, ~lm(baby_weight_g ~ baby_length_cm + ga_weeks, data = .x)),
    head_length_sex_mod = map(train, ~lm(baby_weight_g ~ baby_head_cm * baby_length_cm * baby_sex, data = .x)),
    my_mod = map(train, ~lm(baby_weight_g ~ mom_age + mom_weight_lb + ga_weeks + baby_length_cm + smoke_n, data = .x))
    ) %>% 
  mutate(
    rmse_length_ga = map2_dbl(.x = length_ga_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_head_length_sex = map2_dbl(.x = head_length_sex_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_my_mod = map2_dbl(.x = my_mod, .y = test, ~rmse(model = .x, data = .y))
  )
```

As we did in class, let's compare the models:
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```
Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

## Problem 3
Class notes:
fit linear model of max temp against min temp. Interested in R squared, and then also log(B0*B1) - don't delta method
Download dataset, fit model, make sure you can compute r2 - using broom glance, and the other thing using broom tidy, then get estimated coefficients, then rearrange things so that you have intercept and slope available to take log. Try this with one sample first, THEN figure out how to do with bootstrap.


For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 0∗β̂ 1)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).


